Inattentive responding can induce spurious associations between task behavior and symptom measures

### 2021-04-02 manuscripts/main.tex
push updated manuscript

In the next section, we explicitly test this hypothesis in a large sample of online participants completing a battery of self-report surveys and a behavioral task. Specifically, we measure the correspondence between the common task- and survey-based screening methods, as identified in our literature review, so that the results are informative with respect to the average online psychiatry study.

@@

%In the majority of the papers reviewed above, experimenters relied solely on behavioral screening to exclude participants who engaged in C/IE responding during an experiment. Nevertheless, even with behavioral screening, it is still possible that spurious behavior-symptom correlations might emerge if there is low correspondence between behavioral and survey-based C/IE screening methods (since high correspondence between methods would entail that behavioral screening alone would exclude participants who would have failed survey screening).

@@

%To assess the overall quality of the data, we examined the number of participants flagged by the choice accuracy and infrequency item screening measures (i.e. two most common screening measures identified in our literature review). 
Only 26 participants (7\%) were flagged as exhibiting choice behavior at or below statistically chance levels in the reversal-learning task. In contrast, 85 participants (22\%) endorsed a logically invalid or improbable response on one or more of the infrequency items when completing the self-report symptom measures. This discrepancy in the proportion of participants flagged by each method is consistent with previous research, which found varying levels of sensitivity to C/IE responding across screening methods \cite{desimone2018dirty}. The proportion of participants flagged for C/IE responding was marginally but significantly greater on Mechanical Turk compared to Prolific for both task (MTurk: N=18/186; Prolific: N=8/200; $z=2.22, p=0.026$) and survey data (MTurk: 50/186; Prolific: 35/200; $z = 2.22, p = 0.026$).

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low. To assess this, we visually inspected for each self-report measure the distribution of symptom scores in participants flagged by the infrequency measure vs. those that were not (Figure \ref{fig:distributions}). In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7; Table \ref{tab:validation}). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct.

\begin{table}[!t]
\centering
\small
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lccccccc}
\toprule
& & \multicolumn{2}{c}{Total Score} & \multicolumn{2}{c}{Cronbach's $\alpha$} & \multicolumn{2}{c}{\% Clinical Cutoff} \\
\cmidrule(lr){3-4} \cmidrule(lr){5-6} \cmidrule(lr){7-8}
 Subscale & Skew & Attentive & C/IE & Attentive & C/IE & No Screening & With Screening \\
\midrule
7-up   &  0.81 &    3.87 &   10.15 &    0.84 &    0.84 &     13.0\% &      4.0\% \\
7-down &  0.76 &    4.75 &   10.68 &    0.94 &    0.88 &     17.4\% &      9.3\% \\
GAD-7  &  0.75 &    4.86 &    9.73 &    0.92 &    0.87 &     25.9\% &     17.3\% \\
BIS    &  0.78 &    7.74 &    7.92 &    0.83 &    0.62 &        - &        - \\
BAS    &  0.17 &   15.73 &   16.16 &    0.84 &    0.71 &        - &        - \\
SHAPS  &  0.26 &    8.02 &   10.85 &    0.90 &    0.81 &     17.9\% &     14.6\% \\
PSWQ   &  0.19 &    4.78 &    6.74 &    0.93 &    0.81 &      7.3\% &      7.0\% \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{Descriptive statistics of the self-report symptom measures between attentive and C/IE participants. Skew: the empirical skewness of the distribution of total symptom scores. Total score: the average symptom score between attentive and C/IE participants. Cronbach's $\alpha$: a measure of response consistency, where greater values closer to 1 indicate greater consistency in responses. \% Clinical Cutoff: the percentage of participants reaching threshold for clinical symptomology before and after screening based on the infrequency measure. (The BIS/BAS scales do not have clinical thresholds.)}
\label{tab:validation}
\end{table}

@@

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = 0.67, 95\% \ \text{HDI} = [-0.76, -0.57]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less severe} symptoms on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition).

Finally, we compared the proportion of participants meeting the cutoff for clinical levels of psychopathology before and after excluding participants based on their responses to the infrequency items. Previous studies have found that applying such measures reduced the prevalence of clinical symptomology in online samples towards ground truth rates from epidemiological studies \cite{ophir2020turker}. On the most positively-skewed measures, the fraction of participants reaching clinical levels of symptom endorsement prior to screening was greater than what would be expected (Table \ref{tab:validation}). For example, 13.0\% of participants scored at or above clinical thresholds for (hypo)mania on the 7-up scale in our sample prior to screening, compared with a 12-month prevalence of 5\% in the general population \cite{merikangas2007lifetime, merikangas2012true}, but this rate was reduced to 4.0\% (in line with the population prevalence estimates) after exclusion of C/IE respondents. We observed a similar pattern for both major depressive disorder and anxiety\footnote{Interestingly, compared to previous literature the proportion of participants meeting threshold on the GAD-7 was elevated. We suspect this may reflect elevated rates of state anxiety during the COVID-19 pandemic \cite{yarrington2021impact}, when these data were collected, as well as elevated ground-truth rates of anxiety among the self-selected group of people who choose to work on online labour markets.} (population prevalence estimates of 7\% and 5\% respectively; \cite{kessler2012twelve, lowe2008validation, hinz2017psychometric}).  In line with previous research, we interpret these inflated rates of clinical symptomology in our sample prior to screening as suggestive of C/IE responding \cite{ophir2020turker}.

@@

\begin{figure}[!t]
\includegraphics[width=17cm]{../figures/main_03a.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. Only statistically significant correlations are specified (not corrected for multiple comparisons). Black Xs indicate significant correlations ablated under screening. No Screening = no exclusions; Accuracy Only = exclusions based on chance-level performance in the reversal-learning task; Infrequency Only = exclusions based on invalid or improbable responses to infrequency items; Both Types = exclusions based on the previous two measures. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
\label{fig:correlations}
\end{figure}

@@

These findings suggest that many of the significant behavior-symptom correlations observed without strict participant screening may indeed be spurious correlations driven by C/IE responding. This was particularly the case for self-report instruments that have skewed response distributions resulting from uncommon symptom endorsement on average. Importantly, screening based on task behavior alone did not adequately protect against spurious symptom-behavior correlations in the presence of skewed distributions of symptom endorsement. For instance, consider the 7-down scale, a measure of trait depression: had we not screened participants based on infrequency items, we would have erroneously concluded that there were many significant associations between reversal-learning task performance and self-reported depression. Screening on self-report data allowed us to identify that each of these depression-behavior correlations was likely to be entirely spurious.

One possible objection to this interpretation is that the reduction in significant correlations following self-report screening and rejection was a result of the smaller sample size after removal of C/IE respondents (which comprised over 20\% of the sample). To test this alternative hypothesis, we compared the average (absolute) correlation size before and after infrequency-based exclusions were applied, restricted to correlations that were no longer significant afterward screening (the complete tables of correlations are included in the supplement). On average, we observed a significant reduction in correlation effect sizes ($\Delta \rho = -0.05, t = -2.87, p = 0.012$). following rejections, suggesting that the reduction in significant correlations following screening was unlikely to be driven solely by a reduction in statistical power.

@@

The outputs of the bootstrapping analysis are presented in Figure \ref{fig:bootstrap}. We found that, although estimated correlation magnitudes were independent of sample size (x-axis, left panel), the absolute magnitude of the behavior-symptom correlation increased with the proportion of C/IE participants (different coloured circles, left panel). Crucially, we found false-positive rates for spurious correlations \emph{increased} with increases in sample size in our data (right panel). This runs counter to a common assumption that larger sample sizes are protective against spurious correlations because they serve to mitigate measurement error. Although this assumption is correct for unsystematic measurement error, it no longer holds in the regime of systematic measurement error (where larger sample sizes reduce the variance of estimates, but do not alter their bias). Instead, our results suggest that, holding the proportion of C/IE responding constant, the false-positive rate for behavioral-symptom correlations will become increasingly inflated as the sample size increases.

\begin{figure}[!t]
\includegraphics[width=16cm]{../figures/main_03b.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{False positive rates for spurious correlations \emph{increase} with sample size. \textit{Left:} Bootstrapped Spearman correlations between learning rate asymmetry ($\kappa$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. The grey error bars indicate 95\% bootstrap confidence intervals. The thick dotted lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. \textit{Right:} False positive rates for learning rate asymmetry ($\kappa$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. False positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\kappa$ and 7-down was statistically significant.}
\label{fig:bootstrap}
\end{figure}

@@

Finally, we employed a theory-agnostic modeling approach to better understand how participants suspected of engaging in C/IE responding performed on the probabilistic reversal-learning task. This analysis was motivated by two interrelated aims. First was to better understand why there was low correspondence between screening measures derived from task and self-report data; that is, to determine why participants flagged by the infrequency items were not necessarily more likely to perform at chance. The second was to better understand  the provenance of the particular pattern of behavior-symptom correlations we found.

To characterize participants' choice behavior, we adapted the softmax regression model from \cite{seymour2012serotonin}. This model estimates, for each participant, how much their propensity to choose an option depends on the recent history of trial events (rewarding outcomes, non-rewarding outcomes, and choices from the preceding 5 trials; see Methods). The regression weights for each event, averaged within attentive and careless participants, are presented in Figure \ref{fig:regression}. Comparing attentive participants to those suspected of engaging in C/IE responding, we observed a credible difference (i.e. 95\% highest density intervals excluded zero) only for the $T-1$ weight for previous choice. Attentive participants were more likely to repeat their previous choice (i.e. greater choice hysteresis) than were careless participants.

\begin{figure}[!b]
\includegraphics[width=16cm]{../figures/main_03c.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Softmax regression decision weights in attentive (red) and careless (blue) participants. The weights dictate the extent to which the recent history of rewards, nonrewards, or previous choices influence current choice. $^* p<0.05$}
\label{fig:regression}
\end{figure}

This result may be initially surprising, since one might expect choice hysteresis to result in more perseveration errors following contingency reversals, and since results reported in the previous section indicated that C/IE responding was associated with more perseveration errors. However, choice hysteresis is adaptive in this probabilistic reversal-learning task. Since rewards in the task are probabilistic, ignoring unrewarding outcomes and instead performing the same response once the reward-maximizing response option has been identified is optimal, until the reversal occurs and is identified. Interestingly, participants engaging in C/IE responding were also numerically (though not significantly) less affected by previous outcomes, suggesting that their behavior was not more adaptive, but rather just more random.

This pattern of results, also mirrored in overall reduced outcome sensitivity ($\beta$), explains the pre-screening pattern of correlations with win-stay and lose-shift rates. But what of the learning rate asymmetries? %That learning rate asymmetries ostensibly relate to C/IE responding is puzzling at first given that the softmax regression results show no credible differences in the sensitivity to history of rewarding and non-rewarding outcomes. 
This too can be explained by choice hysteresis. Previous work has established that, when choice hysteresis is not accounted for in reinforcement learning models, it can manifest as positive learning rate asymmetries \cite{Katahira2018-ke}. As such, given that attentive participants show increased repetition, which our reinforcement learning model did not explicitly account for, we should expect to find a negative (positive) correlation between positive (negative) learning rates and C/IE responding. Indeed, this is precisely the pattern of correlations observed above.

To conclude, the theory-agnostic analysis of task behavior revealed that careless participants were less likely to engage in adaptive choice hysteresis and instead showed a pattern of responding more akin to win-stay lose-shift heuristic. This finding helps clarify in part why we observed low correspondence between task and self-report screening measures: on average, careless participants were not engaging in random responding (which would have been detected by the accuracy screening measure) but instead in satisficing behaviors. Most importantly, the results present another hidden danger of C/IE responding. When participants utilize heuristic decision-making, this may manifest in difficult-to-predict forms when computational models are misspecified. Thus, C/IE responding may not only drive spurious correlations but also spurious inferences about what behavior-symptom correlations entail.

@@

One complicating factor for our argument is that C/IE responding may manifest in other ways than simply random responding for both behavioral tasks and self-report surveys. Indeed, in general there are more ways to respond carelessly than to respond attentively to a task or self-report inventory (e.g., random response selection, straight-lining, zig-zagging, acquiescence bias) \cite{curran2016methods}. Indeed, we observed that a win-stay lose-shift pattern emerged as a dominant strategy in the behavioral task among participants suspected of C/IE responding. The specific response strategy a participant adopts is likely to reflect the idiosyncratic integration of multiple perceived benefits (e.g. time saved, effort avoided) and costs (e.g. loss of performance bonuses, risk of detection and pay forfeited). As has been previously documented \cite{desimone2018dirty}, the presence of multiple response strategies makes it clear why certain screening measures are more or less likely to correlate. For example, the inter-item standard deviation and personal reliability measures are both sensitive to statistically random responding, but less sensitive to straight-lining. Most importantly, a diversity of heuristic response strategies highlights the need for many screening measures of C/IE responding, each sensitive to different heuristic strategies.

@@

% offcut material below

Before continuing, we wish to briefly highlight two additional results from our literature review. First, we found that the majority of studies used only a single method to detect C/IE responding. This is worrisome from the perspective of signal detection theory. As has been previously reported \cite{desimone2018dirty}, measures of C/IE responding vary in their sensitivities: some measures will identify only the most careless participants, whereas others will flag many more participants. As such, using only a single measure runs the risk of false positives (in the case of high sensitivity) and false negatives (in the case of low sensitivity). Similarly, screening methods vary in their sensitivity to different patterns of C/IE responding \cite{curran2016methods}. Using only one measure may thus only identify a subset of careless respondents. As has been discussed elsewhere, the use of multiple screening methods is encouraged to more accurately and sensitively identify C/IE responding in a dataset \cite{curran2016methods, barends2019noncompliant}.

Second, we found that the majority of studies that did screen self-report survey data relied on instructed attention checks. This is potentially problematic, since a growing body of research disputes the efficacy of instructed items as attention checks for online studies \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Indeed, instructed attention checks are often discussed on forums for workers of online labour markets\footnote{Several examples of such discussions can be found at the Github repository for this project: \url{github.com/nivlab/sciops}}, and scripts have even been developed with the purpose to highlight words most commonly used in instructed attention checks. Unsurprisingly, then, it has been documented that subsets of participants may actively seek out instructed items in surveys and respond randomly elsewhere \cite{barends2019noncompliant}. In line with best-practice recommendations for online survey research, therefore, we encourage experimenters to use alternate means of screening for C/IE responding for self-report survey data (e.g., infrequency-item attention checks) \cite{huang2015detecting}.

In sum, we found that almost a quarter of participants in our online sample were suspected of C/IE responding when screened using the infrequency-item measure, in contrast to only 7\% of participants flagged when using task accuracy as a screening measure. Importantly, participants suspected of C/IE responding were more likely to appear as symptomatic on self-report measures where symptom endorsement was rare on average. Furthermore, using three independent analyses, we identified evidence that validates the infrequency-item measures as sensitive to C/IE responding. These validation analyses also identified multiple forms of C/IE responding. Though the Cronbach's $\alpha$ result suggested at least some participants were engaging in random responding, the item-response model analysis identified participants who exhibited a preference for the rightmost response options regardless of their semantic content. That C/IE responding in these data manifested in more than one way is consistent with previous reports \cite{curran2016methods}, and further highlights the importance of utilizing multiple screening methods.

It is important to note that these results do not give a complete picture of the likely effects of C/IE responding on symptom-behavior correlations. A full picture would require investigating the effects of C/IE responding over numerous different task designs, symptoms measures, sample sizes, and proportions of careless responders. Nevertheless, the results of this bootstrapping analysis suggest that spurious correlations driven by C/IE responding are likely to be a considerable source of potential error wherever large online samples are collected with asymmetric distributions of symptom endorsement. Indeed, it is also worth highlighting that false positive rates were inflated (well above the expected 5\% level given $\alpha = 0.05$) even with only 5\% of the bootstrapped participant samples engaging in C/IE responding, which further highlights how pernicious this issue is.
% end offcut material

@@

    \item When collecting self-report questionnaire data, include attention-check items that flag participants who may be engaging in C/IE responding. We recommend following best-practice guidelines [citation] in using infrequency-item checks rather than instructed-item checks. Participants flagged by suspicious responses on attention-check items should either be excluded from further analysis, or assessed using sensitivity analyses to ensure that observed full-sample correlations are robust to their exclusion.

@@

    \item In our experience, we have found it instructive to review discussions on public forums for participants of online labour markets (e.g. Reddit, TurkerNation). Doing so helps an experimenter identify what screening methods would-be participants are already aware of and prepared to answer correctly.

@@

To validate the infrequency items as an efficacious measure of C/IE responding, we performed three complimentary analyses. We describe each in turn below.

>>>>

In the next section, we explicitly test this hypothesis in a large sample of online participants completing a battery of self-report surveys and a behavioral task. Specifically, we measure the empirical correspondence between common task- and survey-based screening methods---as identified in our literature review---so that results are informative with respect to typical study designs in online psychiatry research.

@@

@@

To assess the overall quality of the data, we examined the number of participants flagged by the choice accuracy and infrequency item screening measures. Only 26 participants (7\%) were flagged as exhibiting choice behavior at or below statistically chance levels in the reversal-learning task. In contrast, 85 participants (22\%) endorsed a logically invalid or improbable response on one or more of the infrequency items when completing the self-report symptom measures. This discrepancy in the proportion of participants flagged by each method is consistent with previous research, which found varying levels of sensitivity to C/IE responding across screening methods \cite{desimone2018dirty}. The proportion of participants flagged for C/IE responding was marginally but significantly greater on Mechanical Turk compared to Prolific for both task (MTurk: N=18/186; Prolific: N=8/200; $z=2.22, p=0.026$) and survey data (MTurk: 50/186; Prolific: 35/200; $z = 2.22, p = 0.026$).

@@

We hypothesise that spurious behavior-symptom correlations may emerge due to a mean-shift in the average level of symptom endorsement in participants engaging in C/IE responding relative to attentive participants. In turn, a mean-shift is expected to occur when the overall rate of symptom endorsement is low. To assess this, we visually inspected for each self-report measure the distribution of symptom scores in participants flagged by the infrequency measure vs. those that were not (Figure \ref{fig:distributions}). In line with our predictions, the average level of symptom endorsement was noticeably exaggerated in C/IE-responding participants for the symptom measures where symptom scores were most positively-skewed (7-up, 7-down, GAD-7). In contrast, where there was higher rates of symptom endorsement overall, the distributions of symptom scores between the two groups of participants were less noticeably distinct. Permutation testing confirmed that observed mean-shifts in symptom scores for C/IE participants were statistical significant for the majority of the symptom measures (Table \ref{tab:validation}).

\begin{table}[!t]
\centering
\small
\setlength{\tabcolsep}{6pt}
\begin{tabular}{lcccccccc}
\toprule
& & \multicolumn{3}{c}{Total Score} & \multicolumn{2}{c}{Cronbach's $\alpha$} & \multicolumn{2}{c}{\% Clinical Cutoff} \\
\cmidrule(lr){3-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
 Subscale & Skew & Attentive & C/IE & \textit{t}-value & Attentive & C/IE & Before & After \\
\midrule
7-up   &  0.81 &    3.87 &   10.15 &  -13.31^* &    0.84 &    0.84 &     13.0\% &       4.0\% \\
7-down &  0.76 &    4.75 &   10.68 &   -9.99^* &    0.94 &    0.88 &     17.4\% &       9.3\% \\
GAD-7  &  0.75 &    4.86 &    9.73 &   -7.88^* &    0.92 &    0.87 &     25.9\% &      17.3\% \\
BIS    &  0.78 &    7.74 &    7.92 &   -0.54 &    0.83 &    0.62 &        - &             - \\
BAS    &  0.17 &   15.73 &   16.16 &   -0.91 &    0.84 &    0.71 &        - &             - \\
SHAPS  &  0.26 &    8.02 &   10.85 &   -4.04^* &    0.90 &    0.81 &     17.9\% &      14.6\% \\
PSWQ   &  0.19 &    4.78 &    6.74 &   -4.78^* &    0.93 &    0.81 &      7.3\% &       7.0\% \\
\bottomrule
\end{tabular}
\captionsetup{width=0.88\textwidth}
\caption{Descriptive statistics of the self-report symptom measures between attentive and C/IE participants. Skew: the empirical skewness of the distribution of total symptom scores. Total score: the average symptom score between attentive and C/IE participants. * indicates statistical significance. Cronbach's $\alpha$: a measure of response consistency, where greater values closer to 1 indicate greater consistency in responses. \% Clinical Cutoff: the percentage of participants reaching threshold for clinical symptomology before and after screening based on the infrequency measure. (The BIS/BAS scales do not have clinical thresholds.)}
\label{tab:validation}
\end{table}

@@

Second, we quantified the degree to which participants responded to self-report symptom surveys in a stereotyped fashion; that is, we determined if participants exhibited patterns in their responses that were independent of the contents of the survey items. We fit a random-intercept item factor-analysis model \cite{maydeu2006random} to self-report data (see Methods), and for each participant we estimated an intercept parameter that quantified their bias towards using responses on the left or right side of the response scale, regardless of what that response signifies for a particular self-report measure (e.g., low on one symptom scale versus high on another). We observed a credible difference between the average value of this intercept for the two groups ($\Delta \text{intercept} = 0.67, 95\% \ \text{HDI} = [-0.78, -0.55]$), such that C/IE participants were biased towards using the right-half of survey response options. This translates to a tendency to endorse \emph{more severe} symptoms on the 7-up/7-down and GAD-7 scales (where the rightmost options indicate greater frequency of symptoms) but \emph{less severe} symptoms on the SHAPS and BIS (where the rightmost options indicate lower frequency of symptoms) despite these inventories measuring strongly correlated constructs (i.e. depression and anhedonia, anxiety and behavioral inhibition).

Finally, we compared the proportion of participants meeting the cutoff for clinical levels of psychopathology before and after excluding participants based on their responses to the infrequency items. Previous studies have found that applying such measures reduced the prevalence of clinical symptomology in online samples towards ground truth rates from epidemiological studies \cite{ophir2020turker}. On the most positively-skewed measures, the fraction of participants reaching clinical levels of symptom endorsement prior to screening was greater than what would be expected (Table \ref{tab:validation}). For example, 13.0\% of participants scored at or above clinical thresholds for (hypo)mania on the 7-up scale in our sample prior to screening, compared with a 12-month prevalence of 5\% in the general population \cite{merikangas2007lifetime, merikangas2012true}, but this rate was reduced to 4.0\% (in line with the population prevalence estimates) after exclusion of C/IE respondents. We observed a similar pattern for both major depressive disorder and anxiety\footnote{Interestingly, compared to previous literature the proportion of participants meeting threshold on the GAD-7 was elevated. We suspect this may reflect elevated rates of state anxiety during the COVID-19 pandemic \cite{yarrington2021impact}, when these data were collected.} (population prevalence estimates of 7\% and 5\% respectively; \cite{kessler2012twelve, lowe2008validation, hinz2017psychometric}).  In line with previous research, we interpret these inflated rates of clinical symptomology in our sample prior to screening as suggestive of C/IE responding \cite{ophir2020turker}.

@@

\begin{figure}[!t]
\includegraphics[width=17cm]{../figures/main_03a.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Absolute Spearman rank correlations between task behavior (y-axis) and symptom measures (x-axis) under different regimes of data screening and participant exclusions. Only statistically significant correlations are specified (not corrected for multiple comparisons). Black Xs indicate significant correlations ablated under screening. (The unthresholded correlations are included in the Supplement.) No Screening = no exclusions; Accuracy Only = exclusions based on chance-level performance in the reversal-learning task; Infrequency Only = exclusions based on invalid or improbable responses to infrequency items; Both Types = exclusions based on the previous two measures. Acronyms: Acc = choice accuracy; Pts = total points earned; WS = win-stay rate; LS = lose-shift rate; Pers = perseveration errors; $\beta$ = inverse temperature; $\eta_p$ = positive learning rate; $\eta_n$ = negative learning rate; $\kappa$ = learning rate asymmetry.}
\label{fig:correlations}
\end{figure}

@@

These findings suggest that many of the significant behavior-symptom correlations observed without strict participant screening may indeed be spurious correlations driven by C/IE responding. Importantly, screening based on task behavior alone did not adequately protect against spurious symptom-behavior correlations in the presence of skewed distributions of symptom endorsement. For instance, consider the 7-down scale, a measure of trait depression: had we not screened participants based on infrequency items, we would have erroneously concluded that there were many significant associations between reversal-learning task performance and self-reported depression. Screening on self-report data allowed us to identify that each of these depression-behavior correlations was likely to be entirely spurious.

%One possible objection to this interpretation is that the reduction in significant correlations following self-report screening and rejection was a result of the smaller sample size after removal of C/IE respondents (which comprised over 20\% of the sample). To test this alternative hypothesis, we compared the average (absolute) correlation size before and after infrequency-based exclusions were applied, restricted to correlations that were no longer significant afterward screening (the complete tables of correlations are included in the supplement). On average, we observed a significant reduction in correlation effect sizes ($\Delta \rho = -0.05, t = -2.87, p = 0.012$). following rejections, suggesting that the reduction in significant correlations following screening was unlikely to be driven solely by a reduction in statistical power.

One possible objection to this interpretation is that the reduction in significant correlations following- self-report screening and exclusion was a result of the reduced sample size after removal of C/IE respondents (which comprised over 20\% of the sample). To test this alternative hypothesis, we performed the same correlation analysis after removing random subsets of participants, fixing the sample size to that after excluding C/IE respondents. The rate of significant correlations in the bootstrapped samples are reported in Figure S\ref{fig:random}. Briefly, the pattern of significant correlations visually resembled the correlation structure prior to rejections (Figure \ref{fig:correlations}A), and statistical testing confirmed the pattern was more similar to that before screening than after screening using the infrequency measure (t = 262.49, p < 0.001). Thus, the reduction in significant correlations following screening was unlikely to be driven solely by a reduction in statistical power.

@@

The outputs of the bootstrapping analysis are presented in Figure \ref{fig:bootstrap}. We found that, although estimated correlation magnitudes were independent of sample size (x-axis, left panel), the absolute magnitude of the behavior-symptom correlation increased with the proportion of C/IE participants (different coloured circles, left panel). Crucially, we found false-positive rates for spurious correlations \emph{increased} with increases in sample size in our data for all but the smallest rates of C/IE responding (right panel). This runs counter to a common assumption that larger sample sizes are protective against spurious correlations because they serve to mitigate measurement error. Although this assumption is correct for unsystematic measurement error, it no longer holds in the regime of systematic measurement error (where larger sample sizes reduce the variance of estimates, but do not alter their bias). Instead, our results suggest that, holding the proportion of C/IE responding constant, the false-positive rate for behavioral-symptom correlations will become increasingly inflated as the sample size increases.

\begin{figure}[!t]
\includegraphics[width=16cm]{../figures/main_03b.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{False positive rates for spurious correlations \emph{increase} with sample size. \textit{Left:} Bootstrapped Spearman correlations between learning rate asymmetry ($\kappa$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. The grey error bars indicate 95\% bootstrap confidence intervals. The thick dashed lines indicate the threshold for statistical significance for the Spearman correlation at the corresponding sample size. Markers are jittered along the x-axis for legibility. \textit{Right:} False positive rates for learning rate asymmetry ($\kappa$) and depression scores (7-down) as a function of sample size and proportion of C/IE participants. False positive rate was calculated as the proportion of bootstrap samples in which the correlation between $\kappa$ and 7-down was statistically significant.}
\label{fig:bootstrap}
\end{figure}

@@

Finally, we employed a theory-agnostic modeling approach to investigate how C/IE participants on the probabilistic reversal-learning task compared to attentive participants. The motivation for this analysis was to better understand why C/IE responding was inconsistently predicted by chance-level performance, and also correlated with asymmetric learning rates.

To characterize participants' choice behavior, we adapted the softmax regression model from \cite{seymour2012serotonin}. This model estimates, for each participant, how much their choice depends on the recent history of trial events (rewarding outcomes, non-rewarding outcomes, and choices from the preceding 5 trials; see Methods). The regression weights for each event, averaged within attentive and C/IE participants, are presented in Figure \ref{fig:regression}. Comparing attentive to C/IE participants, we observed a credible difference (i.e. 95\% highest density intervals excluded zero) only for the $T-1$ weight for previous choice. That is, attentive participants were more likely to repeat their previous choice (i.e. greater choice hysteresis) than were C/IE participants.

\begin{figure}[!b]
\includegraphics[width=16cm]{../figures/main_03c.png}
\centering
\captionsetup{width=0.88\textwidth}
\caption{Softmax regression decision weights in attentive (red) and C/IE (blue) participants. The weights dictate the extent to which the recent history of rewards, nonrewards, or previous choices influence current choice. Error bars indicate the 95\% highest density interval of each weight. * denotes where the 95\% highest density interval of the difference in weights excludes zero.}
\label{fig:regression}
\end{figure}

This result may be initially surprising, since one might expect choice hysteresis to result in more perseveration errors following contingency reversals, and since we observed in the previous section that C/IE responding was associated with more perseveration errors. However, choice hysteresis is adaptive in this probabilistic reversal-learning task. Because rewards in the task are probabilistic, once the reward-maximizing response option has been identified ignoring an occasional unrewarding outcome and instead performing the same response is optimal (until the next reversal occurs and is identified). Interestingly, participants engaging in C/IE responding were also numerically (though not significantly) less affected by previous outcomes, suggesting that their behavior was not more adaptive, but rather just more random.

This pattern of results also helps explain the pre-screening correlations with asymmetric learning rates. Previous work has established that, when choice hysteresis is not accounted for in reinforcement learning models, it can manifest as positive learning rate asymmetries \cite{Katahira2018-ke, sugawara2021dissociation}. Since C/IE participants showed increased hysteresis, which our reinforcement learning model did not explicitly account for, we should expect to find a negative correlation between learning-rate asymmetries and symptoms before C/IE participants are excluded. Indeed, this is what we observed above.

% This pattern of results, also mirrored in overall reduced outcome sensitivity, explains the pre-screening pattern of correlations with win-stay and lose-shift rates. But what of the learning rate asymmetries? This too can be explained by choice hysteresis. Previous work has established that, when choice hysteresis is not accounted for in reinforcement learning models, it can manifest as positive learning rate asymmetries \cite{Katahira2018-ke}. As such, given that attentive participants show increased repetition, which our reinforcement learning model did not explicitly account for, we should expect to find a negative (positive) correlation between positive (negative) learning rates and C/IE responding. Indeed, this is precisely the pattern of correlations observed above.

In sum, the theory-agnostic analysis of task behavior revealed that C/IE participants exhibited a qualitatively distinct behavioral strategy on the probabilistic reversal-learning task. C/IE participants showed less adaptive choice hysteresis. Moreover, they were numerically (but not significantly) less sensitive to outcome sensitivities. This latter finding helps clarify in part why we observed low correspondence between task and self-report screening measures (that is, C/IE participants were not significantly more likely to respond randomly during the task). These results also present another hidden danger of C/IE responding: qualitatively distinct patterns of behavior under C/IE responding can bias the estimation of parameters of theoretical interest if not properly accounted for.

%To conclude, the theory-agnostic analysis of task behavior revealed that careless participants exhibited less choice hysteresis and instead showed a pattern of responding more akin to win-stay lose-shift heuristic. This finding helps clarify in part why we observed low correspondence between task and self-report screening measures: on average, careless participants were not engaging in random responding (which would have been detected by the accuracy screening measure) but instead in satisficing behaviors. Most importantly, the results present another hidden danger of C/IE responding. When participants utilize heuristic decision-making, this may manifest in difficult-to-predict forms when computational models are misspecified. Thus, C/IE responding may not only drive spurious correlations but also spurious inferences about what behavior-symptom correlations entail.

@@

One complicating factor for our argument is that C/IE responding may manifest in other ways than simply random responding for both behavioral tasks and self-report surveys. Indeed, in general there are more ways to respond carelessly than to respond attentively to a task or self-report inventory (e.g., random response selection, straight-lining, zig-zagging, acquiescence bias) \cite{curran2016methods}. Indeed, we observed qualitatively distinct behavioral strategy of decreased adaptive hysteresis in C/IE participants in the reversal learning task. The specific response strategy a participant adopts is likely to reflect the idiosyncratic integration of multiple perceived benefits (e.g. time saved, effort avoided) and costs (e.g. loss of performance bonuses, risk of detection and pay forfeited). As has been previously documented \cite{desimone2018dirty}, the presence of multiple response strategies makes it clear why certain screening measures are more or less likely to correlate. For example, the inter-item standard deviation and personal reliability measures are both sensitive to statistically random responding, but less sensitive to straight-lining. Most importantly, a diversity of heuristic response strategies highlights the need for many screening measures of C/IE responding, each sensitive to different heuristic strategies.

@@

@@

    \item When collecting self-report questionnaire data, include attention-check items that flag participants who may be engaging in C/IE responding. We recommend following best-practice guidelines in using infrequency-item checks rather than instructed-item checks \cite{barends2019noncompliant, thomas2017validity, hauser2016attentive}. Participants flagged by suspicious responses on attention-check items should either be excluded from further analysis, or assessed using sensitivity analyses to ensure that observed full-sample correlations are robust to their exclusion.

@@

    \item In our experience, we have found it instructive to review discussions on public forums for participants of online labour markets (e.g. Reddit, TurkerNation). Doing so helps an experimenter identify what screening methods would-be participants are already aware of and prepared to answer correctly. (Several examples of workers discussing common attention checks can be found at the Github repository for this project.)

@@

To validate the infrequency items as a sensitive measure of C/IE responding, we performed three complimentary analyses. We describe each in turn below.
